\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-01-23 papers/jss/article.Rnw
Go back to dissing MessagePack and BSON since Jeroen pointed out some limitations in the messagepack R package.  Remove citations to both, and call them "not widely supported".
I think this wording is a fair compromise in that it still makes the
point I wanted to make that they address a shortcoming of text based
JSON for certain classes of applications by providing more efficient
binary encoding, but they make Jeroen's point that they still suck due
to lack of adoption and weak implementations for R packages so
shouldn't really be held up on the same level as JSON.

Add a paragraph break befo

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
% Spelling Standardization:
% Protocol Buffers, not protocol buffers

@@

Given these requirements, how do we safely and efficiently share intermediate results
between different applications, possibly written in different
languages, and possibly running on different computer systems?
In computer programming, \emph{serialization} is the process of
translating data structures, variables, and session state into a
format that can be stored or transmitted and then reconstructed in the
original form later \citep{clinec++}.
% Reverted to my original above, because the replacement below puts me
% to sleep:
%Such systems require reliable and efficient exchange of intermediate
%results between the individual components, using formats that are
%independent of platform, language, operating system or architecture.
Programming
languages such as \proglang{R}, \proglang{Julia}, \proglang{Java}, and \proglang{Python} include built-in
support for serialization, but the default formats 
are usually language specific and thereby lock the user into a single
environment.

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans, and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
implement functions to parse and generate \texttt{JSON} data from \proglang{R}
objects \citep{rjson,RJSONIO,jsonlite}.

A number of binary formats based on \texttt{JSON} have been proposed
that reduce the parsing cost and improve efficiency.  \pkg{MessagePack}
and \pkg{BSON} both have \proglang{R}
interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
data and thus still duplicate field names with each message sent over
the network or stored in a file.  Such formats also lack support for
versioning when data storage needs evolve over time, or when
application logic and requirement changes dictate updates to the
message format.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
% Spelling Standardization:
% Protocol Buffers, not protocol buffers
% large-scale, not large scale

@@

Given these requirements, how do we safely and efficiently share intermediate results
between different applications, possibly written in different
languages, and possibly running on different computer systems?
In computer programming, \emph{serialization} is the process of
translating data structures, variables, and session state into a
format that can be stored or transmitted and then reconstructed in the
original form later \citep{clinec++}.
% Reverted to my original above, because the replacement below puts me
% to sleep:
%Such systems require reliable and efficient exchange of intermediate
%results between the individual components, using formats that are
%independent of platform, language, operating system or architecture.
Programming
languages such as \proglang{R}, \proglang{Julia}, \proglang{Java}, and \proglang{Python} include built-in
support for serialization, but the default formats 
are usually language-specific and thereby lock the user into a single
environment.

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%

A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans, and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
implement functions to parse and generate \texttt{JSON} data from \proglang{R}
objects \citep{rjson,RJSONIO,jsonlite}.
A number of binary formats based on \texttt{JSON} have been proposed
that reduce the parsing cost and improve efficiency, but these formats
are not widely supported.  Furthermore, such formats lack a separate
schema for the serialized data and thus still duplicate field names
with each message sent over the network or stored in a file.
% and still must send duplicate field names
% with each message since there is no separate schema.
% \pkg{MessagePack}
% and \pkg{BSON} both have \proglang{R}
% interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
% data and thus still duplicate field names with each message sent over
% the network or stored in a file.  Such formats also lack support for
% versioning when data storage needs evolve over time, or when
% application logic and requirement changes dictate updates to the
%message format.
